% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file. 

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\usepackage{comment}
%\usepackage{subfigure}
\usepackage[pdftex]{color}
\usepackage{url}
\usepackage{listings}
\usepackage{textcomp}
\usepackage{gnuplot-lua-tikz}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{minted}
\usepackage{bm}
\usepackage{paralist}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{color,soul}
\usepackage{authblk}
\usepackage{multirow}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}

\newcommand{\dz}[1]{{\color{blue}{\it DZ: #1}}}
\newcommand{\rb}[1]{{\color{red}{\it RB: #1}}}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{%FlashMatrix: Parallel, Scalable Data Analysis with Generalized Matrix Operations
}

%for single author (just remove % characters)
%\author{
%{\rm Your N.\ Here}\\
%Your Institution
%\and
%{\rm Second Name}\\
%Second Institution
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
%} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}


\subsection*{Abstract}
FlashMatrix is a matrix-oriented programming framework for general data analysis
with functional programming interface. It scales matrix operations
beyond memory capacity by utilizing solid-state drives (SSDs) in non-uniform
memory architecture (NUMA). To achieve in-memory performance for matrix-based
data analysis tasks, it is essential to move data efficiently from SSDs to
memory as well as from memory to CPU cache. The core of FlashMatrix is to
fuse matrix operations, which provides opportunities of moving data efficiently
in the deeper memory hierarchy.
\dz{Our techniques for efficient data movement: XXX.}
We implement multiple machine learning algorithms in R to benchmark
the performance of FlashMatrix. The execution of the R implementations in
FlashMatrix has performance comparable to optimized C implementations.
When scaling beyond memory capacity on a large parallel machine, the out-of-core
execution of these R implementations in FlashMatrix has performance comparable
to in-memory execution on a billion-scale dataset. Both in-memory and
out-of-core execution significantly outperform the in-memory execution of
Spark MLlib.

\section{Introduction}
\input{intro}

\section{Related Work}
\input{relwork}

\input{design}

\input{eval}

\section{Conclusion}
\input{conclusion}

%\section{Acknowledgments}

%This work is partially supported by NSF ACI-1261715,
%DARPA GRAPHS N66001-14-1-4028 and
%DARPA SIMPLEX program through SPAWAR contract N66001-15-C-4041.

{\footnotesize \bibliographystyle{acm}
\bibliography{fast17}}


%\theendnotes

\end{document}






